---
title: "Community Ecology"
author: "Amelie"
date: "15/03/2022"
output: html_document
---

```{r message = FALSE, warning = FALSE}
library(dplyr)
library(ape)
library(ggtree)
library(vegan) 
library(ggplot2)
library(cowplot)
source ("./Input/theme_A8.txt")
  theme_set(theme_A8())
Dat <- read.csv("./Input/FloristicSurvey.csv", header = TRUE) 
Dat$Location <- recode(Dat$Location, "i" = "Present", "o" = "Absent")
PlDat <- select(Dat, Claytonia_virginiana:maianthenum_racemosum)  #which plants are present in a sample
row.names(PlDat) <- Dat$ï..Quadrate  #sample IDs
```

1. What effect (if any) does garlic mustard have on the plant community? 
  -does the presence of garlic mustard change the species that are present? 
  
```{r}
Pl_binary <- PlDat
Pl_binary[Pl_binary > 0] <- 1  #reformat the data as presence/absence of a plant species in each community
Pl_dist<- dist(Pl_binary, method = 'binary')
Pl_tree <- nj(Pl_dist)
ggtree(Pl_tree, layout = 'rectangular') %<+% Dat + 
  geom_tiplab(aes(colour = Location), size = 3.5) + 
  guides(colour = guide_legend(title = expression(italic("A. petiolata"))))
```

# make your tree look better 

Evidence is weak that presence/absence of A. petiolata affects the species present in a community (weak clustering) 


2. What has a stronger effect on plant community: the presence/absence of garlic mustard or sampling population? 
```{r}
#using abundance data...
Ab_dist <- vegdist(PlDat, method = 'bray', binary = FALSE)
NMDS <- metaMDS(Ab_dist, k = 2)

PDat <- data.frame(NMDS1 = NMDS$points[, 1],
                   NMDS2 = NMDS$points[, 2],
                   SampleID = Dat$ï..Quadrate, 
                   Location = Dat$Location, 
                   Population = as.character(Dat$Population))

P1 <- qplot(x = NMDS1, y = NMDS2, colour = Population, data = PDat)
P2 <- qplot(x = NMDS1, y = NMDS2, colour = Location, data = PDat)
plot_grid(P1, P2)

```

Population appears to have a stronger influence on plant community than the presence or absence of garlic mustard. The communities in the same population are clustered. 