---
output: html_document
---

# Influence of *Alliaria petiolata* on plant community ecology 
#### Amelie Mahrt-Smith, Queen's University, [Github](https://github.com/ameliems99/Metabarcoding)   
    
Garlic mustard, *Alliaria petiolata*, is an invasive species in North America. It was purportedly introduced by European settlers in the 1800s. Current research in the [Colautti Lab](https://www.ecoevogeno.org/research.html) at Queen's University seeks to identify the effect of this species on native soil ecology, and whether the introduced populations of *A. petiolata* are more vigorous than in its native range. To learn more about garlic mustard, visit the [Opinicon Natural History Blog](https://opinicon.wordpress.com/2020/06/09/garlic-mustard-alliaria-petiolata-brassicaceae/).   
   
This analysis seeks to determine the effect of garlic mustard on plant communities, based on a floristic survey at the Queen’s University Biological Station. More specifically, does the presence of garlic mustard affect plant community composition?   
```{r echo = FALSE, message = FALSE, warning = FALSE}
library(dplyr)
library(ape)
library(ggtree)
library(vegan) 
library(ggplot2)
library(cowplot)
source ("./Input/theme_A8.txt")
  theme_set(theme_A8())
```  

```{r}
Dat <- read.csv("./Input/FloristicSurvey.csv", header = TRUE) 
Dat$Location <- recode(Dat$Location, "i" = "Present", "o" = "Absent")  #is A. petiolata present/absent?
PlDat <- select(Dat, Claytonia_virginiana:maianthenum_racemosum) %>%  #keep plants present in sample
  mutate(Alliaria_petiolata = recode(Dat$Location, "Present" = 1, "Absent" = 0))
row.names(PlDat) <- Dat$ï..Quadrate  #sample IDs

Pl_binary <- PlDat
Pl_binary[Pl_binary > 0] <- 1  #reformat data as presence/absence of a plant species in community
Pl_dist<- dist(Pl_binary, method = 'binary')  #distances btwn communities by species composition
Pl_tree <- nj(Pl_dist)  #neighbour-joining tree
ggtree(Pl_tree, layout = 'rectangular') %<+% Dat + 
  geom_tiplab(aes(colour = Location), size = 3.5) + 
  guides(colour = guide_legend(title = expression(italic("A. petiolata"))))
```
      
**Figure 1.** Clustering of plant communities based on the species present.   
   
Garlic mustard has a weak effect on plant community composition. Fig. 1 shows some clustering by presence/absence of *A. petiolata*. This suggests that for groups of plant communities that are broadly similar, the presence of *A. petiolata* can be used to differentiate between them. However, if *A. petiolata* had a biologically relevant impact on plant community composition---i.e., significantly affected the presence/absence of other species---we would expect to see just two distinct clusters of plant communities: an *A. petiolata* present group and an *A. petiolata* absent group.    
    
The floristic survey included abundance of plant species in addition to the community composition, which allows for a more in-depth analysis of the factors affecting plant community structure. For example, what has a stronger effect on plant community: the presence/absence of garlic mustard or the sampling population? 
```{r results = "hide"}
Ab_dist <- vegdist(PlDat, method = 'bray', binary = FALSE)  #distance btwn communities, by abundance
NMDS <- metaMDS(Ab_dist, k = 2)  #non-metric multi-dimensional scaling, 2 dimensions
PDat <- data.frame(NMDS1 = NMDS$points[, 1],
                   NMDS2 = NMDS$points[, 2],
                   SampleID = Dat$ï..Quadrate, 
                   Location = Dat$Location, 
                   Population = as.character(Dat$Population))

P1 <- qplot(x = NMDS1, y = NMDS2, colour = Population, data = PDat) +  #colour by sampling pop
  coord_fixed(ratio = 1.5) +  #keep plots square
  theme(legend.position = "bottom")
P2 <- qplot(x = NMDS1, y = NMDS2, colour = Location, data = PDat) +  #colour by garlic mustard
  guides(colour = guide_legend(title = expression(italic("A. petiolata")))) + 
  coord_fixed(ratio = 1.5) + 
  theme(legend.position = "bottom")
plot_grid(P1, P2, labels = c("a", "b"))  #arrange plots in same figure
```
    
**Figure 2.** (a) The influence of population on the similarity of plant communities. (b) The influence of *A. petiolata* on the similarity of plant communities.   
    
Population appears to have a stronger influence on plant community than the presence or absence of garlic mustard. The communities in the same population are clustered. 